---
title: "Predicting Housing Prices"
author: "Shujing Yi, Yuehui Gong"
date: '2022-10-10'
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---


# Introduction
As Zillow has realized that its housing market predictions are not as accurate as they could be because they do not factor in enough local intelligence in Mecklenburg, this project is to improve Zillow’s housing market predictions for Mecklenburg.

To build a better housing price prediction regression model, our model includes three types of variables: external amenities, internal characteristics of houses from the given housing price dataset, and services from open data source, and spatial factors.


# Setup
```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries
library(tidyverse)
library(raster)
library(tidycensus)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots


# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```



# Data Wrangling

The given housing price dataset includes the housing price and internal characteristics information for each housing point.
It is input as the base dataset for further feather engineering.

```{r read_data}
sale <- 
  st_read("data/studentData.geojson") %>%
  st_transform('ESRI:102286')

glimpse(sale)
```
```{r}
Mecklenberg <- 
  st_read("data/zipcode/zipcode.shp") %>%
  st_transform('ESRI:102286')
```

The following map visualize the distribution of housing points and their prices, which is represented in color.
```{r}
ggplot() +
  geom_sf(data=Mecklenberg,fill = "grey40", color = "grey10",size= 0.7)+
  geom_sf(data = sale, aes(colour = q5(as.numeric(price))), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(sale,"price"),
                   name="Quintile\nBreaks") +
  labs(title="Home Price, Mecklenberg") +
  mapTheme()
```



## External Amenities
The first part of variables our model is using is the external amenities, the proximity to which could add value to houses. To get the data, we are using the open data portal for Mecklenburg. (http://maps.co.mecklenburg.nc.us/openmapping/data.html)

Our model is considering the following amenities and services:
-Public Transportation: bus stop, station
-Public Amenities: park, school
-Potential negative factors: homeless shelter, flood, Landfill



### Bus Stop

```{r}
busstop <- st_read("data/cats_busstops/CATS_BusStops.shp") #%>%
#st_transform(st_crs(tracts20))
ggplot() +
 geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+
  geom_sf(data = busstop,size= 0.3)+
    labs(title = "Bus Stops, Mecklenberg")
  
```
```{r}
sale <-
  sale %>% 
    mutate(bus_nn1 = nn_function(st_coordinates(sale), st_coordinates(busstop), 1))

```

### Station

```{r}
station<- st_read("data/station_merge/station_merge.shp") 
ggplot() +
 geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+
    geom_sf(data = station) +
    labs(title = "Stations, Mecklenberg")

```
```{r}
sale <-
  sale %>% 
    mutate(station_nn1 = nn_function(st_coordinates(sale), st_coordinates(station), 1))

```


### Park

```{r}
park <- st_read("data/park_locations/Park_Locations.shp") %>%
              st_transform(st_crs(Mecklenberg)) 


ggplot() + geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+
  stat_density2d(data = data.frame(st_coordinates(park)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#25CB10", high = "#FA7800", 
                      breaks=c(0.000000003,0.00000003),
                      labels=c("Minimum","Maximum"), name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Parks, Mecklenberg") 

```

Calculate distance to the nearest park as park_nn1:

```{r}
sale <-
  sale %>% 
    mutate(park_nn1 = nn_function(st_coordinates(sale), st_coordinates(park), 1),
           park_nn3 = nn_function(st_coordinates(sale), st_coordinates(park), 3))

```


### School

```{r}
school_public <- st_read("data/cms_schools/CMS_Schools.shp") %>%
              st_transform(st_crs(Mecklenberg)) %>%
  dplyr::select(city, address)

school_charter <- st_read("data/schools_charter/Schools_Charter.shp") %>%
              st_transform(st_crs(Mecklenberg)) %>%
    dplyr::select(city, address)


school<-rbind(school_public,school_charter)

ggplot()+
    geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data=school)+
    labs(title = "Public Schools in Mecklenberg") 
```
```{r}
sale <-
  sale %>% 
    mutate(school_nn1 = nn_function(st_coordinates(sale), st_coordinates(school), 1),
   school_nn2 = nn_function(st_coordinates(sale), st_coordinates(school), 2),   
   school_nn3 = nn_function(st_coordinates(sale), st_coordinates(school), 3))
```


### Homeless Shelter

```{r}
homelessshelter <- st_read("data/homeless_shelters/Homeless_Shelters.shp") %>%
              st_transform(st_crs(Mecklenberg)) 
ggplot() +
  geom_sf(data = homelessshelter, fill = "red") +
  geom_sf(data=st_union(Mecklenberg),fill = "transparent", color = "grey10",size= 0.7)


ggplot() + geom_sf(data=st_union(Mecklenberg), color = "grey10",size= 0.7) +
  stat_density2d(data = data.frame(st_coordinates(homelessshelter)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#25CB10", high = "#FA7800", 
                      breaks=c(0.000000003,0.00000003),
                      labels=c("Minimum","Maximum"), name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Aggravated Assaults, Boston") +
  mapTheme()


sale <-
  sale %>% 
    mutate(homelessshelter_nn1 = nn_function(st_coordinates(sale), st_coordinates(homelessshelter), 1))
```

### Landfill
```{r}
landfill <- st_read("data/landfills/Landfills.shp") %>%
              st_transform(st_crs(Mecklenberg)) 


ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = landfill, fill = "red") +
   labs(title = "Landfill in Mecklenberg") 

```
```{r}
sale <-
  sale %>% 
    mutate(landfill = nn_function(st_coordinates(sale), st_coordinates(st_centroid(landfill)), 1))
```

### Floodplain

```{r}
flood100y <- st_read("data/femaexisting_100yr_floodplain/FEMAExisting_100yr_Floodplain.shp") %>%
              st_transform(st_crs(Mecklenberg)) 
ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = flood100y, fill = "dark blue",color="transparent") +
  labs(title = "100 year Floodplain in Mecklenberg") 
```

```{r}
salegroup <- 
  rbind(
    sale[flood100y,] %>%
      st_drop_geometry() %>%
      left_join(sale) %>%
      st_sf() %>%
      mutate(flood100y = 1),
   sale[flood100y, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(sale) %>%
      st_sf() %>%
      mutate(flood100y = 0))

salegroup <- salegroup[!duplicated(salegroup$musaID), ]  

```

Map the houses within floodplain (flood100y == 1)

```{r}
floodpt<-
  salegroup %>%
  filter(flood100y == 1)

ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = flood100y, fill = "dark blue",color="transparent") +
     geom_sf(data = salegroup,  col = "grey80",
          show.legend = "point", size = .2) +
  geom_sf(data = floodpt, col = "red",
          show.legend = "point", size = .7) +
  labs(title = "Houses within 100 year Floodplain in Mecklenberg") 


```



## Internal Characteristics

### Age

Age calculation
```{r}
salegroup <- 
  salegroup %>%
  mutate(age = 2022 - yearbuilt)
```






##Test


```{r fig.height=10, fig.width=12}

st_drop_geometry(salegroup) %>% 
  dplyr::select(price, age, heatedarea, fullbaths, bedrooms,homelessshelter_nn1,park_nn1,park_nn3,landfill,flood100y,school_nn1,school_nn2,school_nn3,bus_nn1,station_nn1) %>%
  filter(price <= 1000000, age<500, bedrooms<20) %>%
  gather(Variable, Value, -price) %>% 
  
   ggplot(aes(Value, price)) +
     geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
     facet_wrap(~Variable, ncol = 4, scales = "free") +
     labs(title = "Price as a function of continuous variables") +
     plotTheme()

```



```{r fig.height=10, fig.width=12}
st_drop_geometry(salegroup) %>% 
  dplyr::select(price, descbuildi, storyheigh, aheatingty, heatedfuel,actype,extwall,foundation, bldggrade,
                sale_year) %>%
  filter(price <= 1000000) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
     facet_wrap(~Variable, ncol = 4, scales = "free") +
     labs(title = "Price as a function of\ncategorical variables", y = "Mean_Price") +
     plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
numericVars <- 
  select_if(st_drop_geometry(salegroup), is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 
```

##categorical change

### Age

```{r}
s <-salegroup %>% 
  filter(age <= 200) 
histogram(s$age,breaks=40)
```


```{r}
salegroup <- 
  salegroup %>%
  mutate(age.cat = case_when(
                  age >= 0 & age <15  ~ "Up to 15yrs",
                  age >= 15 & age < 50  ~ "15-50yrs",
                  age >= 50 & age < 75  ~ "50-75yrs",                   
                  age >= 75  ~ "75+yrs"),)
```

### Story height
```{r}
glimpse(s$storyheigh)


salegroup$story.cat <- recode(salegroup$storyheigh,"1.0 STORY"= "1story","1 STORY"= "1story", "1.5 STORY" = "2story", "2.0 STORY" = "2story","BI-LEVEL"="2story","SPLIT LEVEL"="2story","2.5 STORY"="3story","3.0 STORY"="3story","CAPE COD"="other","RANCH W/BSMT"="other")

```


## Spatial Factors
### Census Tract
```{r}
census_api_key("a78f5d0c0fd577cd3f16136b5a87d6ecdb62ffba", overwrite = TRUE)

tracts <-  
  get_acs(geography = "tract",
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2020, state="NC",
          county="Mecklenburg County", geometry=TRUE) %>% 
  st_transform('ESRI:102286')

```
```{r}
tracts <- 
  tracts %>%
  dplyr::select( -NAME, -moe) %>%
  spread(variable, estimate) %>%
  rename(TotalPop = B25026_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

tracts <- 
  tracts %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0)) %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)
```

```{r}
ggplot()+
  geom_sf(data = tracts, fill = "transparent", color = "black")+
  geom_sf(data=salegroup, 
          show.legend = "point", size= 0.1) +
  labs(title = "Mecklenburg Census Tracts") +
  theme(plot.title = element_text(size=22))
```
###Intersection

```{r}
salegroup<-st_intersection(salegroup,tracts)

```

```{r}
salegroup<-st_intersection(salegroup,Mecklenberg)
```


# Regression Model

Outer:homelessshelter_nn1,park_nn1,park_nn3,landfill,flood100y,school_nn1,school_nn2,school_nn3,bus_nn1,station_nn1
Census:TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc,MedRent


```{r}
reg1 <- lm(price ~ ., data = st_drop_geometry(salegroup) %>% 
                                 dplyr::select(price, zip,heatedarea, bedrooms, 
                                               age.cat,storyheigh,fullbaths, halfbaths,
                                               numfirepla,foundation,heatedfuel,actype,units,accounttyp, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc))
summary(reg1)
```

# Cross validation

```{r}
inTrain <- createDataPartition(
              y = paste(salegroup$storyheigh.cat, salegroup$heatedfuel, salegroup$actype), 
              p = .60, list = FALSE)
Mecklenberg.training <- salegroup[inTrain,] 
Mecklenberg.test <- salegroup[-inTrain,]  
```

```{r}
reg.training <- lm(price ~ ., data = st_drop_geometry(Mecklenberg.training) %>% 
                                 dplyr::select(price, heatedarea, bedrooms, 
                                               age.cat,story.cat,fullbaths, halfbaths,
                                               numfirepla,foundation,heatedfuel,actype,units,accounttyp, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc))
summary(reg.training)
```


##Test Set 

```{r}
Mecklenberg.test <-
  Mecklenberg.test %>%
  mutate(Regression = "Baseline Regression",
         SalePrice.Predict = predict(reg.training, Mecklenberg.test),
         SalePrice.Error = SalePrice.Predict - price,
         SalePrice.AbsError = abs(SalePrice.Predict - price),
         SalePrice.APE = (abs(SalePrice.Predict - price)) / SalePrice.Predict)%>%
  filter(price < 5000000) 
```

```{r}
ggplot() +
  geom_sf(data=Mecklenberg,fill = "grey40", color = "grey10",size= 0.7)+
  geom_sf(data = Mecklenberg.test, aes(colour = q5(as.numeric(SalePrice.AbsError))), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(sale,"price"),
                   name="Quintile\nBreaks") +
  labs(title="Home Price Absolute Error, Mecklenberg") +
  mapTheme()
```



```{r}
fitControl <- trainControl(method = "cv", number = 100)
set.seed(825)

reg.cv <- 
  train(price ~ ., data = st_drop_geometry(salegroup) %>% 
                                dplyr::select(price, heatedarea, bedrooms, 
                                               age.cat,story.cat,fullbaths, halfbaths,
                                               numfirepla,foundation,heatedfuel,actype,units,accounttyp, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc), 
     method = "lm", trControl = fitControl, na.action = na.pass)

reg.cv
```
```{r}
reg.cv$resample[1:5,]
```


```{r goodness_metrics, message = FALSE, warning = FALSE}
dplyr::select(reg.cv$resample, -Resample) %>%
  gather(metric, value) %>%
  left_join(gather(reg.cv$results[2:4], metric, mean)) %>%
  ggplot(aes(value)) + 
    geom_histogram(bins=35, fill = "#FF006A") +
    facet_wrap(~metric,scales = "free_x") +
    geom_vline(aes(xintercept = mean), colour = "#981FAC", linetype = 3, size = 1.5) +

    labs(x="Goodness of Fit", y="Count", title="CV Goodness of Fit Metrics",
         subtitle = "Across-fold mean reprented as dotted lines")

```

#Spatial Correlation of error


```{r}
coords.test <- st_coordinates(Mecklenberg.test) 

neighborList.test <- knn2nb(knearneigh(coords.test, 5))

spatialWeights.test <- nb2listw(neighborList.test, style="W")

Mecklenberg.test$lagPrice <- lag.listw(spatialWeights.test, Mecklenberg.test$price)
Mecklenberg.test$lagPriceError<-lag.listw(spatialWeights.test, Mecklenberg.test$SalePrice.Error, NAOK = TRUE)

#Mecklenberg.test %>% 
  #mutate(lagPriceError = lag.listw(spatialWeights.test, SalePrice.Error, NAOK = TRUE)) %>%
  #ggplot(aes(lagPriceError, SalePrice.Error))

```
```{r}
ggplot(Mecklenberg.test, aes(x=lagPrice, y=price)) +
  geom_point(colour = "#FA7800") +
  geom_smooth(method = "lm", se = FALSE, colour = "#25CB10") +
  labs(title = "Price as a function of the spatial lag of price",
       caption = "Figure 4.2",
       x = "Spatial lag of price (Mean price of 5 nearest neighbors)",
       y = "Sale Price") +
  plotTheme()
```

```{r}
ggplot(Mecklenberg.test, aes(x=lagPriceError, y=price)) +
  geom_point(colour = "#FA7800") +
  geom_smooth(method = "lm", se = FALSE, colour = "#25CB10") +
  labs(title = "Error as a function of the spatial lag of price",
       caption="Figure 4.3",
       x = "Spatial lag of errors (Mean error of 5 nearest neighbors)",
       y = "Sale Price") +
  plotTheme()
```
#Moran's


```{r}
moranTest <- moran.mc(Mecklenberg.test$SalePrice.Error, 
                      spatialWeights.test, nsim = 999, na.action=na.omit)

ggplot(as.data.frame(moranTest$res[c(1:999)]), aes(moranTest$res[c(1:999)])) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(aes(xintercept = moranTest$statistic), colour = "#FA7800",size=1) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title="Observed and permuted Moran's I",
       subtitle= "Observed Moran's I in orange",
       x="Moran's I",
       y="Count") +
  plotTheme()
```

#Accounting for neighborhood



```{r}
# Errors by neighborhood
left_join(
  st_drop_geometry(Mecklenberg.test) %>%
    group_by(zip) %>%
    summarize(meanPrice = mean(price, na.rm = T)),
  mutate(Mecklenberg.test, predict.fe = 
                        predict(lm(price ~ zip, data = Mecklenberg.test), 
                        Mecklenberg.test)) %>%
    st_drop_geometry %>%
    group_by(zip) %>%
      summarize(meanPrediction = mean(predict.fe))) %>%
      kable() %>% kable_styling()
```


```{r}
reg.nhood <- lm(price ~ ., data = as.data.frame(Mecklenberg.training) %>% 
                                 dplyr::select(zip, price, heatedarea, bedrooms, 
                                               age.cat,story.cat,fullbaths, halfbaths,
                                               numfirepla,foundation,heatedfuel,actype,units,accounttyp, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc))

Mecklenberg.test.nhood <-
  Mecklenberg.test %>%
  mutate(Regression = "Neighborhood Effects",
         SalePrice.Predict = predict(reg.nhood, Mecklenberg.test),
         SalePrice.Error = SalePrice.Predict- price,
         SalePrice.AbsError = abs(SalePrice.Predict- price),
         SalePrice.APE = (abs(SalePrice.Predict- price)) / price)%>%
  filter(price < 5000000)
```


```{r}
summary(reg.nhood)
```

#Accuracy of the neighborhood mode
```{r}
bothRegressions <- 
  rbind(
    dplyr::select(Mecklenberg.test, starts_with("price"), Regression, zip, SalePrice.Predict,SalePrice.Error,SalePrice.AbsError,SalePrice.APE) ,
    dplyr::select(Mecklenberg.test.nhood, starts_with("price"), Regression, zip, SalePrice.Predict, SalePrice.Error,SalePrice.AbsError,SalePrice.APE))   
```

```{r}

st_drop_geometry(bothRegressions) %>% 
   mutate_if(is.numeric, ~ifelse(abs(.) == Inf,NA,.))%>% 
  gather(Variable, Value, -Regression, -zip) %>%
  filter(Variable == "SalePrice.AbsError" | Variable == "SalePrice.APE") %>%
  group_by(Regression, Variable) %>%
    summarize(meanValue = mean(Value, na.rm = T)) %>%
    spread(Variable, meanValue) %>%
    kable()
```
```{r}
bothRegressions %>%
  dplyr::select(SalePrice.Predict, price, Regression) %>%
    ggplot(aes(price, SalePrice.Predict)) +
  geom_point() +
  stat_smooth(aes(price, price), 
             method = "lm", se = FALSE, size = 1, colour="#FA7800") + 
  stat_smooth(aes(SalePrice.Predict, price), 
              method = "lm", se = FALSE, size = 1, colour="#25CB10") +
  facet_wrap(~Regression) +
  labs(title="Predicted sale price as a function of observed price",
       subtitle="Orange line represents a perfect prediction; Green line represents prediction") +
  plotTheme()


```

```{r}



  
st_drop_geometry(bothRegressions) %>%
  group_by(Regression, zip) %>%
  summarize(mean.MAPE = mean(SalePrice.APE, na.rm = T)) %>%
  ungroup() %>% 
  left_join(Mecklenberg) %>%
    st_sf() %>%
    ggplot() + 
      geom_sf(aes(fill = mean.MAPE)) +
      geom_sf(data = bothRegressions, colour = "black", size = .5) +
      facet_wrap(~Regression) +
      scale_fill_gradient(low = palette5[1], high = palette5[5],
                          name = "MAPE") +
      labs(title = "Mean test set MAPE by neighborhood") +
      mapTheme()
```



```{r}
Mecklenberg.training2 <-
  Mecklenberg.training %>%
  mutate(Regression = "Neighborhood Effects",
         SalePrice.Predict = predict(reg.nhood, Mecklenberg.training),
        )
```


```{r}
CHALLENGE<-
  rbind(Mecklenberg.test.nhood[Mecklenberg.test.nhood$toPredict=="CHALLENGE",c("SalePrice.Predict","musaID")],Mecklenberg.training2[Mecklenberg.training2$toPredict=="CHALLENGE",c("musaID","SalePrice.Predict")])%>%
  st_drop_geometry%>%
  rename(prediction = SalePrice.Predict)%>%
subset(select=c(2,1)) 

write.csv(CHALLENGE,"SY_group_prediction.csv", row.names = FALSE)
```

