


```{r setup, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries
library(tidyverse)
library(raster)
library(tidycensus)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots


# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```


## Data Wrangling

See if you can change the chunk options to get rid of the text outputs

```{r read_data}
sale <- 
  st_read("data/studentData.geojson") %>%
  st_transform('ESRI:102286')
```



```{r}
Mecklenberg <- 
  st_read("data/npa/npa.shp") %>%
  st_transform('ESRI:102286')
```

```{r}
ggplot() +
  geom_sf(data=Mecklenberg,fill = "grey40", color = "grey10",size= 0.7)+
  geom_sf(data = sale, aes(colour = q5(as.numeric(price))), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(sale,"price"),
                   name="Quintile\nBreaks") +
  labs(title="Home Price, Mecklenberg") +
  mapTheme()
```




##bike lanes

```{r}
bikelanes <- st_read("data/bikelanes/BikeLanes.shp") %>%
              st_transform(st_crs(Mecklenberg)) 
plot(bikelanes)
```
```{r}
busstop <- st_read("data/cats_busstops/CATS_BusStops.shp") #%>%
#st_transform(st_crs(tracts20))
ggplot() +
  geom_sf(data = busstop) +
  geom_sf(data=st_union(Mecklenberg),fill = "transparent", color = "grey10",size= 0.7)
```
```{r}
sale <-
  sale %>% 
    mutate(bus_nn1 = nn_function(st_coordinates(sale), st_coordinates(busstop), 1))

```

##Station

```{r}
station<- st_read("data/station_merge/station_merge.shp") 
ggplot() +
  geom_sf(data = station) +
  geom_sf(data=st_union(Mecklenberg),fill = "transparent", color = "grey10",size= 0.7)

```
```{r}
sale <-
  sale %>% 
    mutate(station_nn1 = nn_function(st_coordinates(sale), st_coordinates(station), 1))

```



##park
```{r}
park <- st_read("data/park_locations/Park_Locations.shp") %>%
              st_transform(st_crs(Mecklenberg)) 


ggplot() + geom_sf(data=st_union(Mecklenberg), color = "grey10",size= 0.7) +
  stat_density2d(data = data.frame(st_coordinates(park)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#25CB10", high = "#FA7800", 
                      breaks=c(0.000000003,0.00000003),
                      labels=c("Minimum","Maximum"), name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Park, Mecklenberg") +
  mapTheme()
```

Calculate distance to the nearest park as park_nn1:

```{r}
sale <-
  sale %>% 
    mutate(park_nn1 = nn_function(st_coordinates(sale), st_coordinates(park), 1),
           park_nn3 = nn_function(st_coordinates(sale), st_coordinates(park), 3))

```

#Landfill
```{r}
landfill <- st_read("data/landfills/Landfills.shp") %>%
              st_transform(st_crs(Mecklenberg)) 


ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = landfill, fill = "red") +
   labs(title = "Landfill in Mecklenberg") 

```
```{r}
sale <-
  sale %>% 
    mutate(landfill = nn_function(st_coordinates(sale), st_coordinates(st_centroid(landfill)), 1))
```



##School

```{r}
school_public <- st_read("data/cms_schools/CMS_Schools.shp") %>%
              st_transform(st_crs(Mecklenberg)) %>%
  dplyr::select(city, address)

school_charter <- st_read("data/schools_charter/Schools_Charter.shp") %>%
              st_transform(st_crs(Mecklenberg)) %>%
    dplyr::select(city, address)


school<-rbind(school_public,school_charter)

ggplot()+
    geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data=school)+
    labs(title = "Public Schools in Mecklenberg") 
```
```{r}
sale <-
  sale %>% 
    mutate(school_nn1 = nn_function(st_coordinates(sale), st_coordinates(school), 1),
   school_nn2 = nn_function(st_coordinates(sale), st_coordinates(school), 2),   
   school_nn3 = nn_function(st_coordinates(sale), st_coordinates(school), 3))
```


##Homeless Shelter

```{r}
homelessshelter <- st_read("data/homeless_shelters/Homeless_Shelters.shp") %>%
              st_transform(st_crs(Mecklenberg)) 
ggplot() +
  geom_sf(data = homelessshelter, fill = "red") +
  geom_sf(data=st_union(Mecklenberg),fill = "transparent", color = "grey10",size= 0.7)


ggplot() + geom_sf(data=st_union(Mecklenberg), color = "grey10",size= 0.7) +
  stat_density2d(data = data.frame(st_coordinates(homelessshelter)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_gradient(low = "#25CB10", high = "#FA7800", 
                      breaks=c(0.000000003,0.00000003),
                      labels=c("Minimum","Maximum"), name = "Density") +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Aggravated Assaults, Boston") +
  mapTheme()


sale <-
  sale %>% 
    mutate(homelessshelter_nn1 = nn_function(st_coordinates(sale), st_coordinates(homelessshelter), 1))
```



##Floodplain

```{r}
flood100y <- st_read("data/femaexisting_100yr_floodplain/FEMAExisting_100yr_Floodplain.shp") %>%
              st_transform(st_crs(Mecklenberg)) 
ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = flood100y, fill = "dark blue",color="transparent") +
  labs(title = "100 year Floodplain in Mecklenberg") 
```

```{r}
salegroup <- 
  rbind(
    sale[flood100y,] %>%
      st_drop_geometry() %>%
      left_join(sale) %>%
      st_sf() %>%
      mutate(flood100y = 1),
   sale[flood100y, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(sale) %>%
      st_sf() %>%
      mutate(flood100y = 0))
```

Map the houses within floodplain (flood100y == 1)
```{r}
floodpt<-
  salegroup %>%
  filter(flood100y == 1)

ggplot() +
  geom_sf(data=Mecklenberg,fill = "white", color = "grey60",size= 0.5)+  
  geom_sf(data = flood100y, fill = "dark blue",color="transparent") +
     geom_sf(data = salegroup,  col = "grey80",
          show.legend = "point", size = .2) +
  geom_sf(data = floodpt, col = "red",
          show.legend = "point", size = .7) +
  labs(title = "Houses within 100 year Floodplain in Mecklenberg") 


```
Age calculation

```{r}
salegroup <- 
  salegroup %>%
  mutate(age = 2022 - yearbuilt)
```



##Test


```{r fig.height=10, fig.width=12}

st_drop_geometry(salegroup) %>% 
  dplyr::select(price, age, heatedarea, fullbaths, bedrooms,homelessshelter_nn1,park_nn1,park_nn3,landfill,flood100y,school_nn1,school_nn2,school_nn3,bus_nn1,station_nn1) %>%
  filter(price <= 1000000, age<500, bedrooms<20) %>%
  gather(Variable, Value, -price) %>% 
  
   ggplot(aes(Value, price)) +
     geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
     facet_wrap(~Variable, ncol = 4, scales = "free") +
     labs(title = "Price as a function of continuous variables") +
     plotTheme()

```



These variables look weird: Park_nn3, Floodplain, station_nn123,bus,school_nn12

```{r}
table(salegroup$bedrooms)
```


```{r fig.height=10, fig.width=12}
st_drop_geometry(salegroup) %>% 
  dplyr::select(price, descbuildi, storyheigh, aheatingty, heatedfuel,actype,extwall,foundation, bldggrade,
                sale_year) %>%
  filter(price <= 1000000) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
     facet_wrap(~Variable, ncol = 4, scales = "free") +
     labs(title = "Price as a function of\ncategorical variables", y = "Mean_Price") +
     plotTheme() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
numericVars <- 
  select_if(st_drop_geometry(salegroup), is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 
```

##categorical change

### Age

```{r}
s <-salegroup %>% 
  filter(age <= 200) 
histogram(s$age,breaks=40)
```


```{r}
salegroup <- 
  salegroup %>%
  mutate(age.cat = case_when(
                  age >= 0 & age <15  ~ "Up to 15yrs",
                  age >= 15 & age < 50  ~ "15-50yrs",
                  age >= 50 & age < 75  ~ "50-75yrs",                   
                  age >= 75  ~ "75+yrs"),)
```

### Story height
```{r}
glimpse(s$storyheigh)


salegroup$story.cat <- recode(salegroup$storyheigh,"1.0 STORY"= "1story","1 STORY"= "1story", "1.5 STORY" = "2story", "2.0 STORY" = "2story","BI-LEVEL"="2story","SPLIT LEVEL"="2story","2.5 STORY"="3story","3.0 STORY"="3story","CAPE COD"="other","RANCH W/BSMT"="other")

```


## Census Tract
```{r}
census_api_key("a78f5d0c0fd577cd3f16136b5a87d6ecdb62ffba", overwrite = TRUE)

tracts <-  
  get_acs(geography = "tract",
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2020, state="NC",
          county="Mecklenburg County", geometry=TRUE) %>% 
  st_transform('ESRI:102286')

```
```{r}
tracts <- 
  tracts %>%
  dplyr::select( -NAME, -moe) %>%
  spread(variable, estimate) %>%
  rename(TotalPop = B25026_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

tracts <- 
  tracts %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0)) %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)
```

```{r}
ggplot()+
  geom_sf(data = tracts, fill = "transparent", color = "black")+
  geom_sf(data=salegroup, 
          show.legend = "point", size= 0.5) +
  labs(title = "Mecklenburg Census Tracts") +
  theme(plot.title = element_text(size=22))
```
###Intersection

```{r}
salegroup<-st_intersection(salegroup,tracts)
```



##regression model

Outer:homelessshelter_nn1,park_nn1,park_nn3,landfill,flood100y,school_nn1,school_nn2,school_nn3,bus_nn1,station_nn1
Census:TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc,MedRent


```{r}
reg1 <- lm(price ~ ., data = st_drop_geometry(salegroup) %>% 
                                 dplyr::select(price, heatedarea, bedrooms, 
                                               age.cat,story.cat,fullbaths, halfbaths,
                                               numfirepla,foundation, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc))
summary(reg1)
```

```{r}
inTrain <- createDataPartition(
              y = paste(salegroup$storyheigh.cat, salegroup$heatedfuel, salegroup$actype), 
              p = .60, list = FALSE)
Mecklenberg.training <- salegroup[inTrain,] 
Mecklenberg.test <- salegroup[-inTrain,]  
```

```{r}
reg.training <- lm(price ~ ., data = st_drop_geometry(Mecklenberg.training) %>% 
                                 dplyr::select(price, heatedarea, bedrooms, 
                                               age.cat,story.cat,fullbaths, halfbaths,
                                               numfirepla,foundation, park_nn1,homelessshelter_nn1,landfill,flood100y,school_nn1,bus_nn1,station_nn1,bldggrade,
                                               TotalPop,pctWhite,pctBachelors,pctPoverty,MedHHInc))
summary(reg.training)
```

##Test Set 

```{r}
Mecklenberg.test <-
  Mecklenberg.test %>%
  mutate(Regression = "Baseline Regression",
         SalePrice.Predict = predict(reg.training, Mecklenberg.test),
         SalePrice.Error = SalePrice.Predict - price,
         SalePrice.AbsError = abs(SalePrice.Predict - price),
         SalePrice.APE = (abs(SalePrice.Predict - price)) / SalePrice.Predict)%>%
  filter(price < 5000000) 
```

```{r}
ggplot() +
  geom_sf(data=Mecklenberg,fill = "grey40", color = "grey10",size= 0.7)+
  geom_sf(data = Mecklenberg.test, aes(colour = q5(as.numeric(SalePrice.AbsError))), 
          show.legend = "point", size = .75) +
  scale_colour_manual(values = palette5,
                   labels=qBr(sale,"price"),
                   name="Quintile\nBreaks") +
  labs(title="Home Price Absolute Error, Mecklenberg") +
  mapTheme()
```
